### 2024-09-26 manuscript.tex
Multiple improvements

\documentclass[12pt,a4paper]{article}
\usepackage{fontspec}
\usepackage[margin=2cm]{geometry}
\usepackage[citecolor=black]{hyperref}
\usepackage[authoryear,round]{natbib}
\usepackage{doi}
\usepackage{titlesec}
\usepackage{caption}
\DeclareCaptionLabelSeparator{pipe}{ | }
\captionsetup{labelfont={bf,sf},labelsep=pipe}
\usepackage{graphicx}
\usepackage{lineno}
\linenumbers
\setsansfont{Source Sans Pro}
%\setmainfont{Crimson Text}
\setmainfont{CrimsonText}[
	Path=./fonts/,
	Extension=.ttf,
	UprightFont=*-Regular,
	BoldFont=*-Bold,
	ItalicFont=*-Italic,
	BoldItalicFont=*-BoldItalic
]
%\setmonofont[Scale=0.8]{Space Mono}
\setmonofont[Scale=0.8]{SpaceMono}[
	Path=./fonts/,
	Extension=.ttf,
	UprightFont=*-Regular,
	BoldFont=*-Bold,
	ItalicFont=*-Italic,
	BoldItalicFont=*-BoldItalic
]
\hypersetup{
	colorlinks=true,
	urlcolor=blue,
	pdftitle={Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses},
	pdfauthor={Peter Kuma, Frida A.-M. Bender, Adrian J. McDonald and Thorsten Mauritsen}
}
\bibliographystyle{abbrvnat}
\titleformat*{\section}{\large\bfseries\sffamily}
\titleformat*{\subsection}{\normalsize\bfseries\sffamily}
\titlespacing*{\section}{0pt}{12pt}{6pt}
\titlespacing*{\subsection}{0pt}{10pt}{4pt}
%\renewcommand{\bibsection}{\section*{Publications}}
\parskip=6pt
\parindent=0pt

@@

We show that the model underestimates the total cloud fraction by about 10\%,
with overestimation of cloud below 2 km and underestimation of cloud above 2
km. The reanalyses also underestimate the total cloud fraction by about 20\%.
ERA5 overestimates cloud below 1 km but underestimates near-surface cloud or
fog. In addition to lidar data, we compare radiosonde profiles acquired on the
RV \textit{Polarstern} voyages with ICON. Notably, the model exhibits smaller
natural variability than observations, and its lifting condensation level tends
to be higher. This might explain why cloud occurrence is peaking higher in the
model (at 500 m) than in observations (at the surface).

@@

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citep{trenberth2010}, and the SO SW CRE
bias is still positive in eight analysed CMIP6 models analysed by
\cite{schuddeboom2021} over high-latitude SO, whereas over low-latitude SO it
tends to be more neutral or negative in some models. Too much absorbed SW
radiation over the SO was also identified in the GSRM SCREAM
\cite{caldwell2021}. Compensating biases are possible, such as the `too few too
bright' cloud bias, characterised by too small cloud fraction and too large
cloud albedo \citep{wall2017,kuma2020}, previously described by
\cite{webb2001}, \cite{weare2004}, \cite{zhang2005}, \cite{karlsson2008},
\cite{nam2012}, \cite{klein2013}, and \cite{bender2017} in other regions and
models. That is, a model maintains a resonable SW radiation balance by
reflecting too much SW radiation by cloudy sky but has too little cloudy sky
area overall. A study by \cite{konsta2022} showed that this type bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \cite{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \citep{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \cite{fiddes2022} and \cite{fiddes2024}.  \cite{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to the Moderate Resolution Imaging
Spectroradiometer (MODIS)] and overestimates downwelling TOA SW radiation on
the order of 10 Wm$^\mathrm{-2}$ in  [relative to the Clouds and the Earth’s
Radiant Energy System (CERES)] in the highest model resolution run (2.5 km).
\cite{zhao2022} reported a similar SW radiation bias in five analysed CMIP6
models over high-latitude SO and total CF underestimation on the order of 10\%
over the entire 40--60°S SO.

@@

The Modern-Era Retrospective analysis for Research and Applications, Version 2
(MERRA-2) is a reanalysis produced by the Global Modeling and Assimilation
Office at the NASA Goddard Space Flight Center \citep{gelaro2017}.  It uses
version 5.12.4 of the Goddard Earth Observing System (GEOS) atmospheric model
\citep{rienecker2008,molod2015}. The reanalysis output analysed here is
available at a spatial resolution of 0.5° of latitude and 0.625° of longitude,
which is about 56 km in the North--South direction and 35 km in the East--West
direction at 60°S. The number of vertical model levels is 72. Here, we use the
following products: 1-hourly instantaneous 2D single-level diagnostics
(M2I1NXASM) for 2-m temperature and humidity; 3-hourly instantaneous 3D
assimilated meteorological fields (M2I3NVASM) for cloud quantities, pressure,
and temperature; 1-hourly average 2D surface flux diagnostics (M2T1NXFLX) for
precipitation and sea ice; and 1-hourly average 2D radiation diagnostics
(M2T1NXRAD) for radiation quantities \citep{merra2}.

@@

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/lts_dist.pdf}
\caption{
Lower tropospheric stability (LTS) distribution in \textbf{(a)} ERA5 and
\textbf{(b)} MERRA-2 calculated for the 31 voyage tracks and one station from
the highest instantaneous temporal resolution data available. Shown is also the
chosen dividing threshold of 12 K for `stable' and `unstable' conditions.
}
\label{fig:lts}
\end{figure}

\subsection{Partitioning by cyclonic activity and stability}

@@

In addition to the above, we partitioned our data into two subsets by
stability. We determined this by calculating lower tropospheric stability (LTS)
as the difference between the potential temperature at 700 hPa and the surface.
Based on a histogram of LTS in ERA5 and MERRA-2 calculated at all voyage tracks
and stations (Fig.  \ref{fig:lts}), we determined a dividing threshold of 12 K
for relatively unstable (< 12 K) and relatively stable (>= 12 K) conditions.

@@

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{img/cyc_stab_dist.pdf}
\caption{
Geographical distribution of \textbf{(a, b)} cyclonic days and \textbf{(b, d)}
relatively stable (LTS > 12 K) time steps in \textbf{(a, c)} ERA5 in years
2010--2013 (inclusive) and \textbf{(b, d)} ICON in model years 2021--2023 (free
running). Cyclonic days are expressed as a fraction of the number of days with
cyclonic activity, defined as grid points located within a double radius of any
cyclone on a given day (UTC), as identified by CyTRACK.
}
\label{fig:cyclone-stability}
\end{figure}

@@

Here, we briefly describe the results of the cyclonic activity and stability
distribution, which is relavant for the subsequent analaysis, because these
quantities are used for subsetting our dataset. Fig.
\ref{fig:cyclone-stability}a shows a geographical distribution of the fraction
of cyclonic days as determined by the cyclone tracking algorithm applied on the
ERA5 reanalysis and ICON data. As expected the strongest cyclonic activity is
in the high-latitude SO zone, and it is relatively zonally symmetric at all
latitudes.  While both reanalysis and the model agree relatively well,
differences in the strength of the local extremes of occurrence are notable,
especially over the Amundsen Sea, which is more cyclonic in the reanalysis, and
around Cape Adere, which is more cyclonic in ICON. There differences might,
however, stem from the relatively short time periods of comparison (4 year) and
the fact that the model is free running.

Fig. \ref{fig:cyclone-stability}b shows a geographical distribution of
relatively stable and unstable conditions as determined by the LTS.  The
relatively unstable conditions are prevalent in the mid SO (50--65°S), which
might be explained by the relatively cold near-surface air overlying relatively
warm sea surface. Relatively stable conditions are prevalent elsewhere over the
SO. The distribution is also less zonally symmetric than the cyclonic activity.
In the high-latitude SO, the presence of sea ice might have significant
stabilising effect \citep{knight2024}. The ERA5 renalysis is also substantially
more stable than ICON across the whole region.

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, and two reanalyses, MERRA-2 and ERA5 (Fig.
\ref{fig:cloud-occurrence-panel}). We aggregated the sources by calculating the
averages and percentiles of all individual profiles (Fig.
\ref{fig:cloud-occurrence}). The analysis shows that the total cloud fraction
(determined as the fraction of profiles with clouds at any height in the lidar
cloud mask) is underestimated in ICON and reanalyses by about 10\% and 20\%,
respectively.

In particular, ICON overestimates cloud occurrence below 1 km and
underestimates it above, MERRA-2 underestimates cloud occurrence at all
heights, especially near the surface, and ERA5 simulates cloud occurrence
relatively well above 1 km, but strongly underestimates it near the surface.
We note that fog or near-surface clouds are strongly lacking in the reanalyses.
As shown in Fig. \ref{fig:cloud-occurrence-panel}, the biases are relatively
consistent across voyages and longitudes. The ICON results are overall better
matching the observations than the reanalyses.

The ICON comparison is limited by the fact that the model is free-running.
Thus, the vertical profiles are not expected to represent the same weather
conditions as in observations, but long-term statistical comparison is still
possible. Only profiles with the same sea ice conditions (present or not) were
included to avoid comparing sea ice with open sea conditions.

The observations show cloud occurrence peaking at the surface, whereas models
show a higher peak (at about 500 m). The models underestimate total cloud
fraction by 10--20\% and show a strong decrease in cloud occurrence near the
surface, but this is not happening in the observations. ICON and ERA5
overestimate cloud occurrence at the peak (between 0--1 km). Above 1 km, ICON
and MERRA-2 understimate cloud occurrence, but ERA5 is very accurate. The too
high peak in models is partly supported by the LCL distribution, which peaks
higher in the models than in the observations (at the surface), although this
is not very pronounced. The northern zone shows a stronger peak of cloud
occurrence near the surface than the southern zone, and this could be because
higher latitudes have more unstable atmospheric profiles. The northern and
southern zones show similar biases in models as in the general case, but ERA5
does not overestimate the peak in the northern zone (near-surface cloud
occurrence is still strongly understimated). Cyclonic situations have a larger
amount of observed cloudiness, including the peak and total cloud fraction. In
these situations, the models are doing a relatively good job of getting the
vertical profile of cloud occurrence right, but still tend to underestimate
cloud occurrence above 1 km and near the surface. Non-cyclonic situations are
similar to the general case. In stable situations cloud occurrence peaks
strongly at the surface in observations, compared to unstable situations, where
the peak is more obtuse and spread across 0--1 km. In stable situations, the
models are doing a fairly good job, but overestimate cloud occurrence at the
peak below 1 km; above 1 km, they show similar biases as in the general case.
In unstable situations the bias in ICON is very pronounced, with a much
stronger peak at about 500 m, ERA5 is underestimating cloud occurrence below 1
km (especially near the surface), and MERRA-2 even more strongly. In all
situations, even when the models overestimate cloud occurrence at some
altitudes, they always substantially underestimate the total cloud fraction.
ICON can be generally characterised as substantially overestimating cloud
occurrence below 1 km and underestimating above, underestimating the total
cloud fraction, and showing greatest biases in unstable and non-cyclonic
condititions. It also shows a peak of cloud occurrence at higher altitude than
obs (500 m vs. near the surface), and correspondingly, its LCL tends to be also
higher. MERRA-2 can be generally characterised as underestimating cloud
occurrence at nearly all altitudes as well as the total cloud fraction, but
mostly above and below 500 m (the peak at 500 m is well-represented). It
struggles the most in the northern zone and in unstable situations. ERA5 can be
generally characterised as representing cloud occurrence correctly above about
1--1.5 km, overestimating below, but underestimating near-surface cloud
occurrence (0--500 m). The total cloud fraction is strongly understimated in
all situations. It has a tendency towards understimation in the northern zone
and unstable situations, conversely, overestimating in the southern zone and
stable conditions.

@@

When analysed by subsets, the cyclonic subset shows the highest cloud cover,
with 8 oktas occurring on one half of such days (Fig. \ref{fig:cloud-cover}d).
This is not represented by ICON or the reanalyses at all. Interestingly, clear
sky days (0 oktas) also have a local maximum peaking at about 15\% in this
subset.  When we contrast the low and high latitude zones, we see that the high
latitude zone tends to have greater cloud cover, peaking at 8 oktas (Fig.
\ref{fig:cloud-cover}c). The high latitude also has almost no clear sky or
small cloud cover cases (0--4 oktas). ICON and the reanalyses represent at
least this characteristic of the distribution well for 0--3 oktas, but
otherwise show bises similar to the general case. One of the greatest biases is
present in ERA5 in the unstable subset, where it peaks at 3 oktas, whereas the
observations peak at 7 oktas and show negligible cloud cover below 5 oktas.

>>>>

\documentclass[12pt,a4paper]{article}
\usepackage{fontspec}
\usepackage[margin=2cm]{geometry}
\usepackage[citecolor=black]{hyperref}
\usepackage[authoryear,round]{natbib}
\usepackage{doi}
\usepackage{titlesec}
\usepackage{caption}
\DeclareCaptionLabelSeparator{pipe}{ | }
\captionsetup{labelfont={bf,sf},labelsep=pipe}
\usepackage{graphicx}
\usepackage{lineno}
\linenumbers
\setsansfont{Source Sans Pro}
%\setmainfont{Crimson Text}
\setmainfont{CrimsonText}[
	Path=./fonts/,
	Extension=.ttf,
	UprightFont=*-Regular,
	BoldFont=*-Bold,
	ItalicFont=*-Italic,
	BoldItalicFont=*-BoldItalic
]
%\setmonofont[Scale=0.8]{Space Mono}
\setmonofont[Scale=0.8]{SpaceMono}[
	Path=./fonts/,
	Extension=.ttf,
	UprightFont=*-Regular,
	BoldFont=*-Bold,
	ItalicFont=*-Italic,
	BoldItalicFont=*-BoldItalic
]
\hypersetup{
	colorlinks=true,
	urlcolor=blue,
	pdftitle={Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses},
	pdfauthor={Peter Kuma, Frida A.-M. Bender, Adrian J. McDonald and Thorsten Mauritsen}
}
\bibliographystyle{abbrvnat}
\titleformat*{\section}{\large\bfseries\sffamily}
\titleformat*{\subsection}{\normalsize\bfseries\sffamily}
\titlespacing*{\section}{0pt}{14pt}{6pt}
\titlespacing*{\subsection}{0pt}{12pt}{4pt}
%\renewcommand{\bibsection}{\section*{Publications}}
\parskip=6pt
\parindent=0pt

@@

We show that the model underestimates the total cloud fraction (CF) by about 10\%,
with overestimation of cloud below 2 km and underestimation of cloud above 2
km. The reanalyses also underestimate the total CF by about 20\%.
ERA5 overestimates cloud below 1 km but underestimates near-surface cloud or
fog. In addition to lidar data, we compare radiosonde profiles acquired on the
RV \textit{Polarstern} voyages with ICON. Notably, the model exhibits smaller
natural variability than observations, and its lifting condensation level tends
to be higher. This might explain why cloud occurrence is peaking higher in the
model (at 500 m) than in observations (at the surface).

@@

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citep{trenberth2010}, and the SO SW CRE
bias is still positive in eight analysed CMIP6 models analysed by
\cite{schuddeboom2021} over high-latitude SO, whereas over low-latitude SO it
tends to be more neutral or negative in some models. Too much absorbed SW
radiation over the SO was also identified in the GSRM SCREAM
\cite{caldwell2021}. Compensating biases are possible, such as the `too few too
bright' cloud bias, characterised by too small cloud fraction and too large
cloud albedo \citep{wall2017,kuma2020}, previously described by
\cite{webb2001}, \cite{weare2004}, \cite{zhang2005}, \cite{karlsson2008},
\cite{nam2012}, \cite{klein2013}, and \cite{bender2017} in other regions and
models. That is, a model maintains a resonable SW radiation balance by
reflecting too much SW radiation by cloudy sky but has too little cloudy sky
area overall. A study by \cite{konsta2022} showed that this type bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \cite{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \citep{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \cite{fiddes2022} and \cite{fiddes2024}.  \cite{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to the Moderate Resolution Imaging
Spectroradiometer (MODIS)] and overestimates downwelling TOA SW radiation on
the order of 10 Wm$^\mathrm{-2}$ in  [relative to the Clouds and the Earth’s
Radiant Energy System (CERES)] in the highest model resolution run (2.5 km).
\cite{zhao2022} reported a similar SW radiation bias in five analysed CMIP6
models over high-latitude SO and the total cloud fraction underestimation on
the order of 10\% over the entire 40--60°S SO.

@@

The Modern-Era Retrospective analysis for Research and Applications, Version 2
(MERRA-2) is a reanalysis produced by the Global Modeling and Assimilation
Office at the NASA Goddard Space Flight Center \citep{gelaro2017}.  It uses
version 5.12.4 of the Goddard Earth Observing System (GEOS) atmospheric model
\citep{rienecker2008,molod2015}. The reanalysis output analysed here is
available at a spatial resolution of 0.5° of latitude and 0.625° of longitude,
which is about 56 km in the North--South direction and 35 km in the East--West
direction at 60°S. The number of vertical model levels is 72. Here, we use the
following products: 1-hourly instantaneous 2D single-level diagnostics
(M2I1NXASM) for 2-m temperature and humidity; 3-hourly instantaneous 3D
assimilated meteorological fields (M2I3NVASM) for cloud quantities, pressure,
and temperature; 1-hourly average 2D surface flux diagnostics (M2T1NXFLX) for
precipitation; and 1-hourly average 2D radiation diagnostics
(M2T1NXRAD) for radiation quantities \citep{merra2}.

@@

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/lts_dist.pdf}
\caption{
Lower tropospheric stability (LTS) distribution in \textbf{(a)} ERA5 and
\textbf{(b)} MERRA-2 calculated for the 31 voyage tracks and one station from
the highest instantaneous temporal resolution data available. Shown is also the
chosen dividing threshold of 12 K for the relatively stable and unstable
conditions.
}
\label{fig:lts}
\end{figure}

\subsection{Partitioning by cyclonic activity and stability}
\label{sec:cyclone-stability}

@@

In addition to the above, we partitioned our data into two subsets by
stability. We determined this by calculating lower tropospheric stability (LTS)
as the difference between the potential temperature at 700 hPa and the surface.
Based on a histogram of LTS in ERA5 and MERRA-2 calculated at all voyage tracks
and stations (Fig.  \ref{fig:lts}), we determined a dividing threshold of 12 K
for the relatively unstable (< 12 K) and the relatively stable (>= 12 K)
conditions.

@@

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{img/cyc_stab_dist.pdf}
\caption{
Geographical distribution of \textbf{(a, b)} cyclonic days and \textbf{(b, d)}
the relatively stable (LTS > 12 K) time steps in \textbf{(a, c)} ERA5 in years
2010--2013 (inclusive) and \textbf{(b, d)} ICON in model years 2021--2023 (free
running). Cyclonic days are expressed as a fraction of the number of days with
cyclonic activity, defined as grid points located within a double radius of any
cyclone on a given day (UTC), as identified by CyTRACK.
}
\label{fig:cyclone-stability}
\end{figure}

@@

Here, we briefly describe the results of the cyclonic activity and stability
distribution, which is relavant for the subsequent analaysis, because these
conditions are used for subsetting our dataset. Fig.
\ref{fig:cyclone-stability}a shows a geographical distribution of the fraction
of cyclonic days as determined by the cyclone tracking algorithm applied on the
ERA5 reanalysis and ICON data (Section \ref{sec:cyclone-stability}). As
expected, the strongest cyclonic activity is in the high-latitude SO zone, and
it is relatively zonally symmetric at all latitudes.  While both reanalysis and
the model agree relatively well, differences in the strength of the local
extremes of occurrence are notable, especially over the Amundsen Sea, which is
more cyclonic in the reanalysis, and around Cape Adere, which is more cyclonic
in ICON. There differences might, however, stem from the relatively short time
periods of comparison (4 year) and the fact that the model is free running.

Fig. \ref{fig:cyclone-stability}b shows a geographical distribution of the
relatively stable and unstable conditions as determined by the LTS (Section
\ref{sec:cyclone-stability}).  The relatively unstable conditions are prevalent
in the mid SO (50--65°S), which might be explained by the relatively cold
near-surface air overlying relatively warm sea surface. The relatively stable
conditions are prevalent elsewhere over the SO. The distribution is also less
zonally symmetric than the cyclonic activity.  In the high-latitude SO, the
presence of sea ice might have significant stabilising effect
\citep{knight2024}. The ERA5 renalysis is also substantially more stable than
ICON across the whole region.

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, and the two reanalyses, ERA5 and MERRA-2
(Fig.  \ref{fig:cloud-occurrence-panel}). We aggregated the data sources
(voyages, campaigns and stations) by calculating the averages and percentiles
of all individual profiles (Fig.  \ref{fig:cloud-occurrence}). The analysis
shows that the total cloud fraction (determined as the fraction of profiles
with clouds at any height in the lidar cloud mask) is underestimated in ICON
and reanalyses by about 10\% and 20\%, respectively. In particular, ICON
overestimates cloud occurrence below 1 km and underestimates it above, MERRA-2
underestimates cloud occurrence at all heights, especially near the surface,
and ERA5 simulates cloud occurrence relatively well above 1 km, but strongly
underestimates it near the surface.  We note that fog or near-surface clouds
are strongly lacking in the reanalyses.  As shown in Fig.
\ref{fig:cloud-occurrence-panel}, the biases are relatively consistent across
the data sources and longitudes. The ICON results are overall better matching
the observations than the reanalyses.

In the general case (Fig. \ref{fig:cloud-occurrence}a), the observations show
cloud occurrence peaking at the surface, whereas models show a higher peak (at
about 500 m). The models underestimate the total cloud fraction by 10--20\% and
show a strong drop in cloud occurrence near the surface, but this is not
supported by the observations. ICON and ERA5 overestimate cloud occurrence at
the peak (between 0--1 km). Above 1 km, ICON and MERRA-2 understimate cloud
occurrence, but ERA5 is very accurate. The too high peak in models is partly
supported by the LCL distribution, which peaks higher in the models than in the
observations (at the surface), although this is not very pronounced.

When subsetted by low- and high-latitude zones (Fig.
\ref{fig:cloud-occurrence}b, c), we can see that the low-latitude SO zone shows
a stronger peak of cloud occurrence near the surface than the high-latitude SO
zone, and this could be because higher latitudes have more unstable atmospheric
profiles. The low- and high-latitude SO zones show similar biases in models as
in the general case, but ERA5 does not overestimate the peak in the
low-latitude SO zone (near-surface cloud occurrence is still strongly
understimated).

When subsetted by cyclonic and non-cyclonic situations (Fig.
\ref{fig:cloud-occurrence}d, e), we can see that the cyclonic situations have a
larger amount of observed cloudiness, including the peak and total cloud
fraction. In these situations, the models are doing a relatively good job of
getting the vertical profile of cloud occurrence right, but still tend to
underestimate cloud occurrence above 1 km and near the surface. Non-cyclonic
situations are similar to the general case.

When subsetted by the relatively stable and unstable conditions (Fig.
\ref{fig:cloud-occurrence}f, g), as defined in Section
\ref{sec:cyclone-stability}, we can see that in the relatively stable
situations cloud occurrence peaks strongly at the surface in observations,
compared to the relatively unstable situations, where the peak is more obtuse
and spread across 0--1 km.  In the relatively stable situations, the models are
doing a fairly good job, but overestimate cloud occurrence at the peak below 1
km; above 1 km, they show similar biases as in the general case.  In the
relatively unstable situations the bias in ICON is very pronounced, with a much
stronger peak at about 500 m, ERA5 is underestimating cloud occurrence below 1
km (especially near the surface), and MERRA-2 even more strongly.

In all situations, even when the models overestimate cloud occurrence at some
altitudes, they always substantially underestimate the total cloud fraction.
ICON can be generally characterised as substantially overestimating cloud
occurrence below 1 km and underestimating above, underestimating the total
cloud fraction, and showing greatest biases in the relatively unstable and
non-cyclonic condititions. It also shows a peak of cloud occurrence at higher
altitude than obs (500 m vs. near the surface), and correspondingly, its LCL
tends to be also higher. MERRA-2 can be generally characterised as
underestimating cloud occurrence at nearly all altitudes as well as the total
cloud fraction, but mostly above and below 500 m (the peak at 500 m is
well-represented). It struggles the most in the low-latitude SO zone and in the
relatively unstable situations. ERA5 can be generally characterised as
representing cloud occurrence correctly above about 1--1.5 km, overestimating
below, but underestimating near-surface cloud occurrence (0--500 m). The total
cloud fraction is strongly understimated in all situations. It has a tendency
towards understimation in the low-latitude SO zone and the relatively unstable
situations, conversely, overestimating in the high-latitude SO zone and the
relatively stable conditions.

@@

When analysed by subsets, the cyclonic subset shows the highest cloud cover,
with 8 oktas occurring on one half of such days (Fig. \ref{fig:cloud-cover}d).
This is not represented by ICON or the reanalyses at all. Interestingly, clear
sky days (0 oktas) also have a local maximum peaking at about 15\% in this
subset.  When we contrast the low and high latitude zones, we see that the high
latitude zone tends to have greater cloud cover, peaking at 8 oktas (Fig.
\ref{fig:cloud-cover}c). The high latitude also has almost no clear sky or
small cloud cover cases (0--4 oktas). ICON and the reanalyses represent at
least this characteristic of the distribution well for 0--3 oktas, but
otherwise show bises similar to the general case. One of the greatest biases is
present in ERA5 in the relatively unstable subset, where it peaks at 3 oktas, whereas the
observations peak at 7 oktas and show negligible cloud cover below 5 oktas.
